\h1{ Search podutiska array with the maximum/minimum amount }

Here we consider the problem of finding podutiska array with a maximum amount ("maximum subarray problem" in English), as well as some of its variations (including the algorithm for the decision version of this problem online --- described algorithm --- KADR (Iaroslav Tverdokhlib)).


\h2{ problem Statement }

Given an array of numbers $a[1 \ldots n]$. You want to find this segment of this list $a[l \ldots r]$ that the sum of \bf{maximum}:

$$ \max_{ 1 \le l \le r \le n } \sum_{i=l}^{r} a[i]. $$

For example, if all numbers of the array $a[]$ would be non-negative, then the answer could be the entire array. The solution is not trivial, when the array may contain both positive and negative numbers.

It is clear that the problem of finding \bf{minimal} podutiska --- essentially the same, just change the signs of all numbers to the opposite.


\h2{ Algorithm 1 }

Here we consider the algorithm is almost obvious. (Next we will consider another algorithm, which is a little harder to figure out, but it is even shorter.)

\h3{ algorithm Description }

The algorithm is very simple.

We introduce for convenience \bf{symbol}: $s[i] = \sum_{j=1}^{i} a[j]$. I.e. the array $s[i]$ is the array of partial sums of an array $a[]$. Also, place the value $s[0] = 0$.

Now let \bf{sort out} the index $r = 1 \ldots n$, and learn for each current value $r$ to quickly find the optimal $l$, which gives the maximum amount in current segment $[l; r]$.

Formally this means that we need for the current $r$ to find $l$ (not greater than $r$) to the value of $s[r] - s[l-1]$. After trivial transformations we get what we need to find in the array $s[]$ minimum on the interval $[0;r-1]$.

From here we immediately get an algorithm for solving: we'll just have to store where in the array s $[]$ is the current minimum. Using this minimum, we have $O(1)$ find the current optimal index $l$, and the transition from the current index $r$ to the next we simply update the minimum.

Obviously, this algorithm runs in $O(n)$ and asymptotically optimal.

\h3{ Implementation }

For the realization we wouldn't even need to explicitly store the array of partial sums $s[]$ --- we will require only the current item.

The implementation is given in 0-indexed arrays rather than 1-numbering, as described above.

Let us give first a solution that finds just the numerical answer, not finding the indices of the corresponding segment:

\code
int ans = a[0],
sum = 0,
min_sum = 0;
for (int r=0; r<n; r++) {
sum += a[r];
ans = max (ans, sum - min_sum);
min_sum = min (min_sum, sum);
}
\endcode

Now we present the complete solution which is in parallel with the numerical solution finds the boundaries of the corresponding segment:

\code
int ans = a[0],
ans_l = 0,
ans_r = 0,
sum = 0,
min_sum = 0,
min_pos = -1;
for (int r=0; r<n; r++) {
sum += a[r];

int cur = sum - min_sum;
if (cur > ans) {
ans = cur;
ans_l = min_pos + 1;
ans_r = r;
}

if (sum < min_sum) {
min_sum = sum;
min_pos = r;
}
}
\endcode


\h2{ Algorithm 2 }

Here we consider another algorithm. Its a little tricky to understand, but it is more elegant than the above, is slightly shorter. This algorithm was proposed by J. Kadan (Jay Kadane) in 1984

\h3{ algorithm Description }

Himself \bf{algorithm} as follows. We will go through the array and store in some variable $s$ current partial sum. If in some point of $s$ is negative, we simply assign $s=0$. It is argued that the maximum of all values of the variable $s$ that happened during the work, and will be the answer to the problem.

\bf{Prove} this algorithm.

Indeed, consider the rst time when the sum $s$ becomes negative. This means that, starting with zero partial sums, we eventually came to a negative partial sum --- so, this whole array of prefix and any suffix have negative amounts. Therefore, from all the prefix of the array can not be any good: it can give only a negative addition to the answer.

But it is not enough to prove the algorithm. In the algorithm we are, in fact, limit the answer to only those segments that begin immediately after the ground, happened when $s<0$.

But, in fact, consider an arbitrary interval $[l;r]$, and $l$ is not in such a "critical" position (i.e., $l > p+1$, where $p$ --- the last position where $s<0$). Since the last critical position is strictly earlier than $l-1$, then $a[p+1 \ldots l-1]$ is nonnegative. This means that by sliding $l$ in position $p+1$, we increase the response or, in extreme cases, not changing it.

Anyway, it turns out that indeed when the answer can be limited to only segments that begin immediately after positions, which were $s<0$. This proves the correctness of the algorithm.

\h3{ Implementation }

As in algorithm 1, we present first a simplified implementation that only looks for a numeric answer, unable to find the boundaries of the corresponding segment:

\code
int ans = a[0],
sum = 0;
for (int r=0; r<n; r++) {
sum += a[r];
ans = max (ans, sum);
sum = max (sum, 0);
}
\endcode

A complete solution to the maintenance of indices of boundaries of the corresponding segment:

\code
int ans = a[0],
ans_l = 0,
ans_r = 0,
sum = 0,
minus_pos = -1;
for (int r=0; r<n; r++) {
sum += a[r];

if (sum > ans) {
ans = sum;
ans_l = minus_pos + 1;
ans_r = r;
}

if (sum < 0) {
sum = 0;
minus_pos = r;
}
}
\endcode


\h2{ Related problems }

\h3{ finding a maximum/minimum podutiska with constraints }

If the problem says on the desired segment $[l;r]$ are imposed additional restrictions (e.g., length $r-l+1$ of the line segment must be outside the prescribed range), then the algorithm is likely easily generalized to these cases --- either way, the task will still be to find minimum in the array $s[]$ given additional restrictions.

\h3{ two-Dimensional case of the problem: finding the maximum/minimum sub-matrices }

Described in this article, the task naturally generalizes to higher dimensions. For example, in the two-dimensional case it turns into a search for such sub-matrices $[l_1 \ldots r_1; l_2 \ldots r_2]$ of the matrix that has the maximum sum of numbers in it.

From the above solution for one-dimensional case \bf{easy to get} the solution for $O(n^3)$: consider $l_1$ and $r_1$, and calculate the amounts array from $l_1$ by $r_1$ in each row of the matrix; we came to a one-dimensional problem of the search of index $l_2$ and $r_2$ in the array, which can be solved in linear time.

\bf{faster} algorithms for solving this task, although well-known, but they are not much faster than $O(n^3)$, and it is very complex (so complex that the hidden constant many of them are inferior to the trivial algorithm when all reasonable limits). Apparently, the best known algorithms work for $O \left( n^3 \frac{ \log^3 \log n }{ \log^2 n} \right)$ (2007 T. Chan, "More algorithms for all-pairs shortest paths in weighted graphs").

This algorithm Chan and many other results in this area actually describe \bf{multiplication} matrices (where matrix multiplication is implied modified multiplication instead of addition is used the least, and instead of multiplying --- add). The fact that the problem of finding the submatrix with the largest sum is reduced to the problem of finding shortest paths between all pairs of vertices, a task which, in turn --- this boils down to matrix multiplication.

\h3{ Search podutiska with maximum/minimum average amount }

This task lies in the fact that it is necessary to find a segment $[l;r]$, so the mean value was maximum:

$$ \max_{l \le r} \frac{ 1 }{ r-l+1 } \sum_{i=l}^{r} a[i]. $$

Of course, if desired the segment $[l;r]$ by the condition imposed no other conditions, then the solution will always be a segment of length $1$ in the point-maximum of the array. The task only makes sense if there are \bf{additional constraints} (for example, the length of the corresponding segment is limited from below).

In such a case is applicable \bf{standard method} when working with tasks about the average value: let's pick up the desired maximum average \bf{binary search}.

For this we need to learn how to solve such a subproblem: given a number $x$, and need to check if there is any segment of this list in the array $a[]$ (of course, satisfies all the additional constraints in the problem), in which the average value is greater than $x$.

To solve this sub-problem, subtract $x$ from each element of the array $a[]$. Then our subproblem actually turns into this: Yes or no in the given array segment of this list a positive amount. And we already know how to solve it.

Thus, we get the solution for simpatico $O (T(n) \log W)$, where $W$ --- the required accuracy, $T(n)$ --- the time of solving the subproblems for an array of length $n$ (which may vary depending upon the particular impose additional restrictions).

\h3{ solution of the problem online }

The problem statement is: given an array of $n$ integers and an integer $L$. Receive requests of the form $(l,r)$, and in response to the request is required to find a segment of this list of the segment $[l;r]$ of length at least $L$ with the maximum possible arithmetic mean.

The algorithm for solving this problem is quite complicated. The author of the algorithm --- KADR (Iaroslav Tverdokhlib) --- \href=http://e-maxx.ru/forum/viewtopic.php?id=410{described this algorithm in his message on the forum}.
